Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2025-02-05 paper/main.tex
Updates

\subsection{Connection to Score Based Methods}
\begin{itemize}
	\item The process of iteratively adding and removing edges in very similar to
		Greedy Equivalence Search (GES).
	\item But not decomposible like scoring methods. 
	\item But Because they make only local changes, can get stuck in local minimas.
	\item Both approach looks for modifications globally.
	\item The sum of unexplained effect provides an absolute measure of model fit 
		unlike scoring methods.
\end{itemize}

>>>>

\subsection{Connection to Score Based Methods}

Our procedure closely resembles certain score-based automated causal discovery
methods, such as Greedy Equivalence Search (GES), where the algorithm
iteratively adds or removes edges that improve the scoring metric the most.
Similarly, a greedy version of our approach would involve adding an edge
between the pair of variables with the highest unexplained correlation. While
both methods aim to find modifications that most improve the model at each
step, a key distinction lies in the nature of the evaluation criteria: unlike
standard scoring metrics, our measure of association is not decomposable. A key
property of scoring metrics is that they should be decomposible, i.e., they can
be expressed as the sum of scores of nodes given its parents. This means adding
or removing an edge affects only a localized part of the model. In contrast,
our measure of association is global, meaning that modifying an edge in one
part of the model can influence association values elsewhere.

Another major difference is in the interpretability of the evaluation metric.
Most scoring metrics are based on log-likelihood with a penalty for model
complexity. These score metrics allow for relative comparisons between models
but do not provide an absolute measure of model fit. That is, they indicate
which model is better for a given dataset but do not reveal how well the model
explains the data in an absolute sense. In contrast, summing our measure of
association across all variables provides a direct fitness measure—this sum
approaches zero when the model perfectly accounts for all observed correlations
in the data. This property enables an absolute assessment of model quality,
rather than just a relative comparison between models.
